---
title: "Homework 17"
author: "Ziyao Yang"
date: "March, 31, 2023"
format: pdf
geometry: 
  - top=30mm
  - left=20mm
editor: visual
---

# \[A\] Calculate E\[X1X2\].

```{r}
M <- matrix(c(0.2,0.1,0.3,0.3,0.05,0.05), nrow = 2, byrow = TRUE)
X1 <- c(1, 2)
X2 <- c(1, 2, 3)
EX1X2 <- sum(X1 %*% t(X2) * M)
EX1X2
```

# \[A\] Calculate the conditional distributions of X1 \| X2 and X2 \| X1

```{r}
# conditional distribution of X1 given X2
PX1_given_X2 <- t(t(M) / colSums(M))
PX1_given_X2
# conditional distribution of X2 given X1
PX2_given_X1 <- M / rowSums(M)
PX2_given_X1

cat("\n")

#conditional distribution of X1 given X2
for (i in 1:2) {
  for (j in 1:3) {
    cat(paste0("Pr(X1 = ", i, " | X2 = ", j, ") = ", M[i, j], " / ", colSums(M)[j], " = ", PX1_given_X2[i, j], "\n"))
  }
}

cat("\n")

#conditional distribution of X2 given X1
for (i in 1:2) {
  for (j in 1:3) {
    cat(paste0("Pr(X2 = ", j, " | X1 = ", i, ") = ", M[i, j], " / ", rowSums(M)[i], " = ", PX2_given_X1[i, j], "\n"))
  }
}
```

# \[C\] Use Gibbs sampling to simulate a Markov Chain of length m = 10000 that starts at X1 = 1, X2 = 1 and evolves according to the conditional distributions of X1 \| X2 and X2 \| X1. Use set.seed(440) in your simulation.

```{r}
set.seed(440)
m = 10000
X1 = 1
X2 = 1

# initialize matrix to store Markov chain
chain <- matrix(0, nrow = m, ncol = 2)
chain[1, ] <- c(X1, X2)
#Gibbs sampling with conditional distribution
for (i in 2:m) {
  # sample X1 given X2
  X1 <- sample(1:2, size = 1, prob = PX1_given_X2[,X2])
  
  # sample X2 given X1
  X2 <- sample(1:3, size = 1, prob = PX2_given_X1[X1,])
  
  # store current values in chain
  chain[i, ] <- c(X1, X2)
}
```

# \[C\] Estimate Pr(X1 = 2, X2 = 1) using your simulated Gibbs samples from the previous part, and then compare it with the theoretical value.

```{r}
# estimate Pr(X1 = 2, X2 = 1)
count <- sum(chain[, 1] == 2 & chain[, 2] == 1)
est_prob <- count / m
cat(paste0("Estimated Pr(X1 = 2, X2 = 1) = ", est_prob, "\n"))
cat(paste0("Theoretical Pr(X1 = 2, X2 = 1) = ", M[2,1], "\n"))
```

# \[C\] Estimate E\[X1X2\] using your simulated Gibbs samples from the previous part, and then compare it with the theoretical value.

```{r}
# estimate E[X1X2]
est_EX1X2 <- mean(chain[, 1] * chain[, 2])
cat(paste0("Estimated E[X1X2] = ", est_EX1X2, "\n"))
cat(paste0("Theoretical E[X1X2] = ", EX1X2, "\n"))
```
